---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---
#1. Ex 8.3
 #we are training and testing dataset
```{r}

library(pacman)

p_load(mdsr, NHANES, tidyverse, Amelia, rpart)

data(NHANES)
head(NHANES)
help("NHANES")
```

Note that there is a lot of missing data in the NHANES dataset.  Start by removing columns with a high percentage of missing values.  They remove row where there is a missing value of the target variable.

```{r}
missmap(NHANES, main = "Missing values vs observed")
```


# number of missing values in each column

```{r}
sapply(NHANES, function(x) sum(is.na(x)))
```


# number of unique values in each column

```{r}
sapply(NHANES, function(x) length(unique(x)))
```


```{r}
NHANES %>% count()

NHANES %>%
  select(everything()) %>%  # replace to your needs
  summarise_all(funs(sum(is.na(.))/10000))
  
```

Remove an columns with a rate of missing that is too high.

```{r}
NHANES2 <- NHANES[, colMeans(is.na(NHANES)) < 0.5] 

NHANES2 %>%
  select(everything()) %>%  # replace to your needs
  summarise_all(funs(sum(is.na(.))/10000))
```

Remove the ID variable.

```{r}
NHANES2 <- NHANES2 %>% select(-ID)
NHANES2 <- NHANES2 %>% select(SleepTrouble, everything())
NHANES2
```

```{r}
NHANES2$SleepTrouble[1:10]

as.numeric(NHANES2$SleepTrouble[1:10])

as.numeric(NHANES2$SleepTrouble[1:10]) - 1

NHANES2$SleepTrouble <- as.numeric(NHANES2$SleepTrouble) - 1

NHANES2$SleepTrouble[1:10]

str(NHANES2)
```

```{r}
missmap(NHANES2, main = "Missing values vs observed")
```

# number of missing values in each column

```{r}
sapply(NHANES2, function(x) sum(is.na(x)))
```


# number of unique values in each column

```{r}
sapply(NHANES2, function(x) length(unique(x)))
```

Remove all rows of the dataset where SleepTrouble is missing.

```{r}
NHANES3 <- NHANES2 %>% drop_na(SleepTrouble)
NHANES3
```

```{r}
missmap(NHANES3, main = "Missing values vs observed")
```

# number of missing values in each column

```{r}
sapply(NHANES3, function(x) sum(is.na(x)))
```


# number of unique values in each column

```{r}
sapply(NHANES3, function(x) length(unique(x)))
```
##KNN-Model
```{r}
library(tidyverse)
library(RNHANES)
library(class)
library(ggplot2)
DEMO_F = nhanes_load_data("DEMO_F", "2009-2010") %>%
  select(SEQN, RIDAGEYR)
BMX_F = nhanes_load_data("BMX_F", "2009-2010") %>% 
  select(SEQN, BMXBMI, BMXWT)
HDL_F =  nhanes_load_data("HDL_F", "2009-2010") %>% 
  select(SEQN, LBDHDD)
GLU_F = nhanes_load_data("GLU_F", "2009-2010") %>% 
  select(SEQN, LBXGLU, LBXIN)
DIQ_F = nhanes_load_data("DIQ_F", "2009-2010") %>% 
  select(SEQN, DIQ010)
dtx = left_join(DEMO_F, HDL_F) %>% 
  left_join(GLU_F) %>% 
  left_join(BMX_F) %>% 
  left_join(DIQ_F)

dat = dtx %>% 
  filter(!is.na(BMXBMI), !is.na(LBDHDD), !is.na(LBXGLU), !is.na(LBXIN),RIDAGEYR >= 40, DIQ010 %in% c(1, 2)) %>% 
  transmute(SEQN, Age = RIDAGEYR, BMI = BMXBMI, Cholest = LBDHDD, Glucose = LBXGLU, Insuline = LBXIN, Weight = BMXWT, SleepTrouble = DIQ010) %>% 
  mutate(sleep_trouble = recode_factor(SleepTrouble, 
                           `1` = "Yes", 
                           `2` = "No"))

#ggplot(dat, aes(Glucose, Weight, color = sleep_trouble)) +
  #geom_point(alpha = 0.7, size = 2) 
#ggplot(dat, aes(Glucose, BMI, color = sleep trouble)) +
 # geom_point(alpha = 0.7, size = 2)
summary(dat)
normalize <- function (i) {
  (i - min(i))/(max(i) - min(i))
}

norm_dat <- dat %>% 
  select(Age, BMI, Cholest, Glucose, Insuline, Weight) %>% 
  lapply(., normalize) %>% 
  as.data.frame()
summary(norm_dat)
dat_samp <- sample(2, nrow(dat), replace=TRUE, prob=c(0.67, 0.33))
dat_training <- norm_dat[dat_samp==1, 1:6]
dat_test <- norm_dat[dat_samp==2, 1:6]
dat_target_group <- dat[dat_samp==1, 8]
dat_test_group <- dat[dat_samp==2, 8]
```
```{r}
#KNN model
dat_pred <- knn(train = dat_training, test = dat_test, cl = dat_target_group, k=3)
dat_pred
summary(dat_pred)
summary(dat_test_group)
tab <- table(dat_pred, dat_test_group)
tab
```
```{r}
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
accuracy(tab)
```


##2.SMS spam Filtering Analysis
```{r}
sms_raw <- read.csv("sms_spam.csv", stringsAsFactors = FALSE)

sms_raw

# examine the structure of the sms data
str(sms_raw)

# convert spam/ham to factor.
sms_raw$type <- factor(sms_raw$type)

# examine the type variable more carefully
str(sms_raw$type)
table(sms_raw$type)

# build a corpus using the text mining (tm) package
library(tm)
sms_corpus <- VCorpus(VectorSource(sms_raw$text))

# examine the sms corpus
print(sms_corpus)
inspect(sms_corpus[1:2])

as.character(sms_corpus[[1]])
lapply(sms_corpus[1:2], as.character)
```
```{r}
# clean up the corpus using tm_map()
sms_corpus_clean <- tm_map(sms_corpus, content_transformer(tolower))

# show the difference between sms_corpus and corpus_clean
as.character(sms_corpus[[1]])
as.character(sms_corpus_clean[[1]])

sms_corpus_clean <- tm_map(sms_corpus_clean, removeNumbers) # remove numbers
sms_corpus_clean <- tm_map(sms_corpus_clean, removeWords, stopwords()) # remove stop words
sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation) # remove punctuation

# tip: create a custom function to replace (rather than remove) punctuation
removePunctuation("hello...world")
replacePunctuation <- function(x) { gsub("[[:punct:]]+", " ", x) }
replacePunctuation("hello...world")

# illustration of word stemming
library(SnowballC)
wordStem(c("learn", "learned", "learning", "learns"))

sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)

sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace) # eliminate unneeded whitespace

# examine the final clean corpus
lapply(sms_corpus[1:3], as.character)
lapply(sms_corpus_clean[1:3], as.character)

# create a document-term sparse matrix
sms_dtm <- DocumentTermMatrix(sms_corpus_clean)

# alternative solution: create a document-term sparse matrix directly from the SMS corpus
sms_dtm2 <- DocumentTermMatrix(sms_corpus, control = list(
  tolower = TRUE,
  removeNumbers = TRUE,
  stopwords = TRUE,
  removePunctuation = TRUE,
  stemming = TRUE
))
```
```{r}
# alternative solution: using custom stop words function ensures identical result
sms_dtm3 <- DocumentTermMatrix(sms_corpus, control = list(
  tolower = TRUE,
  removeNumbers = TRUE,
  stopwords = function(x) { removeWords(x, stopwords()) },
  removePunctuation = TRUE,
  stemming = TRUE
))

# compare the result
sms_dtm
sms_dtm2
sms_dtm3

# creating training and test datasets
sms_dtm_train <- sms_dtm[1:4169, ]
sms_dtm_test  <- sms_dtm[4170:5559, ]

# also save the labels
sms_train_labels <- sms_raw[1:4169, ]$type
sms_test_labels  <- sms_raw[4170:5559, ]$type

# check that the proportion of spam is similar
prop.table(table(sms_train_labels))
prop.table(table(sms_test_labels))

# word cloud visualization
library(wordcloud)
wordcloud(sms_corpus_clean, min.freq = 50, random.order = FALSE)

# subset the training data into spam and ham groups
spam <- subset(sms_raw, type == "spam")
ham  <- subset(sms_raw, type == "ham")

#wordcloud(spam$text, max.words = 40, scale = c(3, 0.5))
#wordcloud(ham$text, max.words = 40, scale = c(3, 0.5))

sms_dtm_freq_train <- removeSparseTerms(sms_dtm_train, 0.999)
sms_dtm_freq_train

# indicator features for frequent words
findFreqTerms(sms_dtm_train, 5)

# save frequently-appearing terms to a character vector
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
```
```{r}
# create DTMs with only the frequent terms
sms_dtm_freq_train <- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]

# convert counts to a factor
convert_counts <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

# apply() convert_counts() to columns of train/test data
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test  <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)

## Step 3: Training a model on the data ----
library(e1071)
sms_classifier <- naiveBayes(sms_train, sms_train_labels)

## Step 4: Evaluating model performance ----
sms_test_pred <- predict(sms_classifier, sms_test)

head(sms_test_pred)

library(gmodels)
CrossTable(sms_test_pred, sms_test_labels,
           prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE,
           dnn = c('predicted', 'actual'))

## Step 5: Improving model performance ----
sms_classifier2 <- naiveBayes(sms_train, sms_train_labels, laplace = 1)
sms_test_pred2 <- predict(sms_classifier2, sms_test)
CrossTable(sms_test_pred2, sms_test_labels,
           prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE,
           dnn = c('predicted', 'actual'))

```